A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-03-19 paper/submission/ZW10157-response-2.tex
edits to new bib entries
small grammer edits

Sundry proofreading corrections listed by section.

I.

paragraph 3
NSGA-II is mentioned but not cited.

II. A.

paragraph 4
By calling the selector --> Calling the selector

@@

III. A.

paragraph 2
search algorithm to get stuck --> search algorithm getting
stuck

III. B.

paragraph 3
checking for idle WORKER --> checking for idle WORKERs

Fig. 3 caption
$r_k$ --> $r_1$

Fig. 3 caption
The work $W_j$ are beam dynamics simulation within OPAL.
--> $W_j$ performs beam dynamics simulations within OPAL.

III. C.

paragraph 1
shown in Listing 2 --> shown in Listing 1

>>>>

{\bf Our response:} {\color{blue} }

Sundry proofreading corrections listed by section.
I. paragraph 3: NSGA-II is mentioned but not cited.

{\bf Our response:} {\color{blue} The citation has been added.}

II. A. paragraph 4: By calling the selector --$>$ Calling the selector

{\bf Our response:} {\color{blue} This correction is in the new manuscript on line xxx.}

@@

{\bf Our response:} {\color{blue} The reference to Cornell was added.}

III. A: paragraph 2
search algorithm to get stuck --> search algorithm getting
stuck

{\bf Our response:} {\color{blue} This correction is in the new manuscript on line xxx.}

III. B. paragraph 3
checking for idle WORKER --> checking for idle WORKERs

{\bf Our response:} {\color{blue} This correction is in the new manuscript on line xxx.}

Fig. 3 caption
$r_k$ --> $r_1$
The work $W_j$ are beam dynamics simulation within OPAL.
--> $W_j$ performs beam dynamics simulations within OPAL.

{\bf Our response:} {\color{blue} The figure caption has been corrected.}

III. C. paragraph 1
shown in Listing 2 --> shown in Listing 1

### 2019-03-19 paper/submission/prab.tex
starting to add appendix
small grammer edits

A modular multi-objective software framework was developed (see
 Fig.~\ref{fig:framenetwork}) where the core functionality of the optimizer is decoupled from
 the ``beam dynamics'' but fully integrated in the OPAL framework. 
To that end, we use a master/slave mechanism where a master process governs a
 set of slave processes given some computational tasks (beam dynamics simulation) to complete.
This separation allows easy interchange of optimization algorithms, forward
  solvers and optimization problems.
A ``pilot'' coordinates all efforts between the optimization algorithm and the
  beam dynamics task. In the following sections, we will also use the notion of ``forward solver'' to indicate the beam dynamics task.
This forms a robust and general framework for massively parallel
  multi-objective optimization.
Currently the framework offers one concrete optimization algorithm, an
  evolutionary algorithm employing a \textsc{NSGA-II} selector \cite{pisa}.
Normally, simulation based approaches are plagued by the trade off between
  level of detail and time to solution.
This problem is addressed later in Section~(\ref{awa:subsection:test}) by using forward solvers with different time and
  resolution complexity.

@@

Since there already exist plenty of implementations of evolutionary algorithms,
  it was decided to incorporate the PISA library \cite{pisa} into our
  framework.
One of the advantages of PISA is that it separates variator from selector,
  rendering the library expandable and configurable.
Implementing a variator was enough to use PISA in our framework and
  retain access to all available PISA selectors.
As shown in Fig.~\ref{fig:varsel}, the selector is in charge of ordering a
  set of $d$-dimensional vectors and selecting the $k$ fittest individuals
  currently in the population.
The performance of a selector depends on the number of objectives and the
  surface of the search space.
So far, the NSGA-II selector \cite{dpam:02} has been used and exhibits satisfactory
  convergence performance.

The task of the variator is to generate offspring and ensure diversity in the
  population.
The variator can start generating offspring once the fitness of every
  individual of the population has been evaluated.
This explicit synchronization point defines an obvious bottleneck for parallel
  implementations of evolutionary algorithms.
In the worst case, some MPI processes are taking a long time to compute the
  fitness of the last individual in the pool of individuals to evaluate.
During this time all other resources are idle and wait for the result of
  this one individual in order to continue to generate and evaluate offspring.
To counteract this effect, the selector is already called when two individuals
  have finished evaluating their fitness, lifting the boundaries between
  generations and evaluating the performance of individuals.
New offspring will be generated and MPI processes can immediately return to
  work on the next fitness evaluation.
By calling the selector more frequently (already after two offspring
  individuals have been evaluated) results in better populations since bad
  solutions are rejected earlier.
On the other hand, calling the selector more often is computationally more
  expensive. Note this capability is also present in GPT's~\cite{gpt} optimization system.

@@

The framework presented in \cite{lems:09} resembles our implementation the
  most, aside from their tight coupling with an evolutionary algorithm
  optimization strategy.
The authors propose a plug-in based framework employing an island
  parallelization model, where multiple populations are evaluated concurrently
  and independently up to a point where some number of individuals of the
  population are exchanged.
This is especially useful to prevent the search algorithm to get stuck in
  a local minimum.
A set of default plug-ins for genetic operators, selectors and other
  components of the algorithms are provided by their framework.
User-based plug-ins can be incorporated into the framework by implementing a
  simple set of functions.

@@

The basic assumption in simulation-based optimization is that a
  call to an expensive simulation software component present in the
  constraints or objectives is needed.
The framework is divided in three exchangeable components, as shown in
  Fig.~\ref{fig:opt-framework-layout}, to encapsulate the major behavioral
  patterns of the framework.
%
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{opt-framework-layout}
  \caption{Schematic view of messages passed within the network between the
    three roles.
  The dashed cyan path describes a request (job $j_1$) sent from $O_i$ to the
  \textsc{Pilot} being handled by $W_j$. Subsequently the result $r_k$ is
  returned to the requesting \textsc{Optimizer} ($O_i$). The work $W_j$ are beam dynamics 
  simulation within OPAL.}
  \label{fig:opt-framework-layout}
\end{figure}

@@

Every available MPI process will take up one of the three available roles (see
  Fig.~\ref{fig:framenetwork}):  one process acts as \textsc{Pilot}, the
  remaining processes are divided amongst \textsc{Worker} and
  \textsc{Optimizer} roles.
Both, the \textsc{Worker} and the \textsc{Optimizer} can consist of multiple
  MPI processes to exploit parallelism.
As shown in Fig.~\ref{fig:opt-framework-layout}, the \textsc{Pilot} is used
  to coordinate all ``information requests'' between the \textsc{Optimizer}
  and the \textsc{Worker}.
An information request is a job that consists of a set of design variables
  (e.g.~the genes of an individual) and a type of information it requests
  (e.g.~function evaluation or derivative).
The \textsc{Pilot} keeps checking for idle \textsc{Worker} and assigns jobs
  in the queue to any free \textsc{Worker}.
Once the \textsc{Worker} has computed and evaluated the request its results
  are returned to the \textsc{Optimizer} that originally requested the
  information.

@@

% AA: this is too much for Phys Rev. AB  
% By abiding to the API from Listing 3 it becomes trivial to add new forward
% solvers to the framework (see Appendix A).

@@

>>>>

A modular multi-objective software framework was developed (see
 Fig.~\ref{fig:framenetwork}) where the core functionality of the optimizer is decoupled from
 the ``beam dynamics'' but fully integrated in the OPAL framework. 
To that end, we use a master/slave mechanism where a master process governs a
 set of slave processes given some computational tasks (beam dynamics simulation) to complete.
This separation allows easy interchange of optimization algorithms, forward
  solvers and optimization problems.
A ``pilot'' coordinates all efforts between the optimization algorithm and the
  beam dynamics task. In the following sections, we will also use the notion of ``forward solver'' to indicate the beam dynamics task.
This forms a robust and general framework for massively parallel
  multi-objective optimization.
Currently the framework offers one concrete optimization algorithm, an
  evolutionary algorithm employing a \textsc{NSGA-II} selector~\cite{dpam:02,pisa}.
Normally, simulation based approaches are plagued by the trade off between
  level of detail and time to solution.
This problem is addressed later in Section~(\ref{awa:subsection:test}) by using forward solvers with different time and
  resolution complexity.

@@

Since there already exist plenty of implementations of evolutionary algorithms,
  it was decided to incorporate the PISA library \cite{pisa} into our
  framework.
One of the advantages of PISA is that it separates variator from selector,
  rendering the library expandable and configurable.
Implementing a variator was enough to use PISA in our framework and
  retain access to all available PISA selectors.
As shown in Fig.~\ref{fig:varsel}, the selector is in charge of ordering a
  set of $d$-dimensional vectors and selecting the $k$ fittest individuals
  currently in the population.
The performance of a selector depends on the number of objectives and the
  surface of the search space.
So far, the NSGA-II selector~\cite{dpam:02} has been used and exhibits satisfactory
  convergence performance.

The task of the variator is to generate offspring and ensure diversity in the
  population.
The variator can start generating offspring once the fitness of every
  individual of the population has been evaluated.
This explicit synchronization point defines an obvious bottleneck for parallel
  implementations of evolutionary algorithms.
In the worst case, some MPI processes are taking a long time to compute the
  fitness of the last individual in the pool of individuals to evaluate.
During this time all other resources are idle and wait for the result of
  this one individual in order to continue to generate and evaluate offspring.
To counteract this effect, the selector is already called when two individuals
  have finished evaluating their fitness, lifting the boundaries between
  generations and evaluating the performance of individuals.
New offspring will be generated and MPI processes can immediately return to
  work on the next fitness evaluation.
Calling the selector more frequently (already after two offspring
  individuals have been evaluated) results in better populations since bad
  solutions are rejected earlier.
On the other hand, calling the selector more often is computationally more
  expensive. Note this capability is also present in Cornell's~\cite{bazarov05} 
  and GPT's~\cite{gpt} optimization system.

@@

The framework presented in \cite{lems:09} resembles our implementation the
  most, aside from their tight coupling with an evolutionary algorithm
  optimization strategy.
The authors propose a plug-in based framework employing an island
  parallelization model, where multiple populations are evaluated concurrently
  and independently up to a point where some number of individuals of the
  population are exchanged.
This is especially useful to prevent the search algorithm getting stuck in
  a local minimum.
A set of default plug-ins for genetic operators, selectors and other
  components of the algorithms are provided by their framework.
User-based plug-ins can be incorporated into the framework by implementing a
  simple set of functions.

@@

The basic assumption in simulation-based optimization is that a
  call to an expensive simulation software component present in the
  constraints or objectives is needed.
The framework is divided in three exchangeable components, as shown in
  Fig.~\ref{fig:opt-framework-layout}, to encapsulate the major behavioral
  patterns of the framework.
%
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{opt-framework-layout}
  \caption{Schematic view of messages passed within the network between the
    three roles.
  The dashed cyan path describes a request (job $j_1$) sent from $O_i$ to the
  \textsc{Pilot} being handled by $W_j$. Subsequently the result $r_1$ is
  returned to the requesting \textsc{Optimizer} ($O_i$). $W_j$ are beam dynamics 
  simulations within OPAL.}
  \label{fig:opt-framework-layout}
\end{figure}

@@

Every available MPI process will take up one of the three available roles (see
  Fig.~\ref{fig:framenetwork}):  one process acts as \textsc{Pilot}, the
  remaining processes are divided amongst \textsc{Worker} and
  \textsc{Optimizer} roles.
Both, the \textsc{Worker} and the \textsc{Optimizer} can consist of multiple
  MPI processes to exploit parallelism.
As shown in Fig.~\ref{fig:opt-framework-layout}, the \textsc{Pilot} is used
  to coordinate all ``information requests'' between the \textsc{Optimizer}
  and the \textsc{Worker}.
An information request is a job that consists of a set of design variables
  (e.g.~the genes of an individual) and a type of information it requests
  (e.g.~function evaluation or derivative).
The \textsc{Pilot} keeps checking for idle \textsc{Worker}s and assigns jobs
  in the queue to any free \textsc{Worker}.
Once the \textsc{Worker} has computed and evaluated the request its results
  are returned to the \textsc{Optimizer} that originally requested the
  information.

@@

@@

\appendix

\section{Code Implementation}

\section{Example Optimization File}

This examples serves to show how a user can 
set up a multiobjective optimization run 
within an OPAL input file.

The code that follows was used to perfrom ex-1 in Section~\ref{sec:experiments}.
